---
title: "transmission_metrics"
author: "Paul Taconet"
date: "24/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) #set t
```

```{r packages}
library(ggplot2)
library(scales)
library(sf)
library(DBI)
library(RSQLite)
library(dplyr)
library(tidyr)
library(purrr)
library(rlang)
library(patchwork)
library(plotly)
library(htmlwidgets)
#library(kable)
library(kableExtra)
library(mapview)
library(stringr)
library(emo)
```

```{r db_connect}
# connect to database
path_to_db <- "data/react_db/react_db.gpkg"
react_gpkg <- DBI::dbConnect(RSQLite::SQLite(),dbname = path_to_db)

# open tables
trmetrics_entomo <- dbReadTable(react_gpkg, 'trmetrics_entomo')
entomo_csh_metadata_l1 <- dbReadTable(react_gpkg, 'entomo_csh_metadata_l1')
recensement_villages_l1 <- st_read(path_to_db,"recensement_villages_l1", quiet=T) %>%
  filter(!is.na(intervention)) %>%
  dplyr::select("codevillage","codepays","nomvillage","population","intervention","date_debut_interv","date_fin_interv","X","Y")

# # join to trmetrics_entomo the tables related to the HLC dates and locations + villages
# trmetrics_entomo <- trmetrics_entomo %>%
#   left_join(entomo_csh_metadata_l1[c("idpointdecapture","nummission","codevillage","pointdecapture","codepays","date_capture","X","Y")], by = "idpointdecapture") %>%
#   left_join(recensement_villages_l1[c("codevillage","intervention","date_debut_interv","date_fin_interv")], by = "codevillage") %>%
#   dplyr::select(-fid)

# get the number of human - nigth
idpointdecaptures <- dbReadTable(react_gpkg, 'entomo_csh_metadata_l1') %>%
  dplyr::select(idpointdecapture,nummission,codevillage,pointdecapture,codepays,date_capture)
csh_exterior <- data.frame(idpointdecaptures, postedecapture = "e")
csh_interior <- data.frame(idpointdecaptures, postedecapture = "i")
csh_all <- rbind(csh_exterior,csh_interior) %>%
  left_join(recensement_villages_l1) %>%
  mutate(phase_interv = ifelse(as.Date(date_capture) < as.Date(date_debut_interv),"pre-intervention","post-intervention")) %>%
  mutate(phase_interv = ifelse((is.na(phase_interv) & codepays=="BF" & as.Date(date_capture) < as.Date("2017-08-17")),"pre-intervention",phase_interv)) %>%
  mutate(phase_interv = ifelse((is.na(phase_interv) & codepays=="CI" & as.Date(date_capture) < as.Date("2017-09-01")),"pre-intervention",phase_interv)) %>%
  mutate(phase_interv = ifelse((is.na(phase_interv) & codepays=="BF" & as.Date(date_capture) > as.Date("2017-08-17")),"post-intervention",phase_interv)) %>%
  mutate(phase_interv = ifelse((is.na(phase_interv) & codepays=="CI" & as.Date(date_capture) > as.Date("2017-09-01")),"post-intervention",phase_interv))


# get mean date for each entomo mission in each country
mean_date_by_mission <- entomo_csh_metadata_l1 %>%
  group_by(codepays,nummission) %>%
  summarise(date = mean(as.Date(date_capture))) %>%
  as_tibble()

# get start and end date of intervention for each country
dates_intervention <- recensement_villages_l1 %>%
  group_by(codepays) %>%
  summarise(start_intervention_date=min(as.Date(date_debut_interv),na.rm = T),end_intervention_date=max(as.Date(date_fin_interv),na.rm = T))
  
```

```{r rainbow_plots}
#source("https://raw.githubusercontent.com/RainCloudPlots/RainCloudPlots/master/tutorial_R/R_rainclouds.R")
source("https://raw.githubusercontent.com/RainCloudPlots/RainCloudPlots/master/tutorial_R/summarySE.R")
#source("https://raw.githubusercontent.com/RainCloudPlots/RainCloudPlots/master/tutorial_R/simulateData.R")
library(cowplot)
library(forcats)

#see https://wellcomeopenresearch.org/articles/4-63
geom_flat_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                             position = "dodge", trim = TRUE, scale = "area",
                             show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomFlatViolin,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      trim = trim,
      scale = scale,
      ...
    )
  )
}

#' @rdname ggplot2-ggproto
#' @format NULL
#' @usage NULL
#' @export
GeomFlatViolin <-
  ggproto("GeomFlatViolin", Geom,
    setup_data = function(data, params) {
      data$width <- data$width %||%
        params$width %||% (resolution(data$x, FALSE) * 0.9)

      # ymin, ymax, xmin, and xmax define the bounding rectangle for each group
      data %>%
        group_by(group) %>%
        mutate(
          ymin = min(y),
          ymax = max(y),
          xmin = x,
          xmax = x + width / 2
        )
    },

    draw_group = function(data, panel_scales, coord) {
      # Find the points for the line to go all the way around
      data <- transform(data,
        xminv = x,
        xmaxv = x + violinwidth * (xmax - x)
      )

      # Make sure it's sorted properly to draw the outline
      newdata <- rbind(
        plyr::arrange(transform(data, x = xminv), y),
        plyr::arrange(transform(data, x = xmaxv), -y)
      )

      # Close the polygon: set first and last point the same
      # Needed for coord_polar and such
      newdata <- rbind(newdata, newdata[1, ])

      ggplot2:::ggname("geom_flat_violin", GeomPolygon$draw_panel(newdata, panel_scales, coord))
    },

    draw_key = draw_key_polygon,

    default_aes = aes(
      weight = 1, colour = "grey20", fill = "white", size = 0.5,
      alpha = NA, linetype = "solid"
    ),

    libraryd_aes = c("x", "y")
  )


a <- trmetrics_entomo %>%
    filter(genre == "Anopheles") %>%
  group_by(idpointdecapture, postedecapture) %>%
  summarise(ma = sum(ma)) %>%
  right_join(csh_all) %>%
  mutate(ma = ifelse(is.na(ma), 0, ma)) %>%
  left_join(mean_date_by_mission) %>%
  as_tibble()



p9 <- ggplot(a %>% filter(codepays=="BF"),aes(x = forcats::fct_rev(phase_interv), y = ma, fill = forcats::fct_rev(phase_interv), colour = forcats::fct_rev(phase_interv)))+
  geom_flat_violin(position = position_nudge(x = .25, y = -1), adjust =2, trim = TRUE)+
  geom_point(position = position_jitter(width = .15), size = .25)+
  geom_boxplot(aes(x = forcats::fct_rev(phase_interv), y = ma),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +
  ylab('Score')+xlab('Group')+theme_cowplot()+guides(fill = FALSE, colour = FALSE) +
  facet_wrap(~intervention)+
  scale_colour_brewer(palette = "Dark2")+
  scale_fill_brewer(palette = "Dark2")+
  scale_y_continuous(limits = c(-1,NA)) +
  ggtitle("Figure R9: Complex Raincloud Plots with Facet Wrap")

p9


```

```{r functions}

fun_agg_data_entomo <- function(trmetrics_entomo, df_n_obs, ...){
  
  group_var <- enquos(...)
  
  df_metric <- trmetrics_entomo %>%
    filter(genre == "Anopheles") %>%
    group_by(!!! group_var) %>%
    summarise(val = sum(ma)) %>%
    left_join(df_n_obs) %>%
    mutate(val = round(val/n_obs,2)) %>%
    left_join(mean_date_by_mission) %>%
    dplyr::select(-n_obs) %>%
    as_tibble()
  
  return(df_metric)

}

fun_plot_ts_entomo <- function(df, indicator_code, aes_color, facet = NULL, title, log2 = FALSE){
  
  aes_color <- enquo(aes_color)
  facet <- enquo(facet)

    th_plot <- ggplot(df, aes(x = date, y = val)) + 
    geom_line(aes(color = !! aes_color), size = 0.5, linetype = "dashed") +
    geom_point(aes(color = !! aes_color), size = 2) +
    #scale_color_manual(values = c("#00AFBB", "#E7B800")) +
    theme_minimal() +
    labs(y = indicator_code) + 
    scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
    #ylim(-0.01,NA) +
    geom_vline(xintercept = as.numeric(mean(dates_intervention$start_intervention_date)),colour = "red", size = 0.3) +
    #annotate(geom="text",label="start of interventions",x=mean(dates_intervention$start_intervention_date)-10, y=mean(ma_b_country$val), colour = "red", angle = 90, size = 3.5) +
    #annotate(geom="text",label="Vertical red line = start of vector control interventions", colour = "red", size = 3.5, y=max(df$val), x=min(df$date)+160) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle(title)
  
  if(log2){
    th_plot <- th_plot + scale_y_continuous(trans = 'log2', limits = c(-0.2,NA), labels = scales::number_format(accuracy = 0.01)) + labs(y = paste0(indicator_code," (log2 scale)"))
  }
   if(!quo_is_null(facet)){
     th_plot <- th_plot + facet_grid(rows = vars(!!facet))
   }

      return(th_plot)
}


```

Click a village to get additional information

```{r map }
areas <- st_read(path_to_db,"contexte_frontieresreact", quiet=T)
mapview(areas, alpha.regions = 0.1, aplha = 1, layer.name = "Study areas") + mapview(recensement_villages_l1, zcol = "intervention", cex = 4, alpha.regions = 0.9, aplha = 1, layer.name = "Villages")
```

##  Ma : Human biting rate 

Average number of bites by Anopheles mosquitoes received per human per night

### By study area

Average number of bites by Anopheles mosquitoes received per human per night, by study area

`r emo::ji("warning")` : **y axis is log2 scaled**

```{r ma_b_country}
# by country

df_n_obs <- csh_all %>%
  group_by(codepays,nummission) %>%
  summarise(n_obs = n())

ma <- fun_agg_data_entomo(trmetrics_entomo, df_n_obs, codepays, nummission)

plot <- fun_plot_ts_entomo(df = ma, 
                           indicator_code = "ma", 
                           aes_color = codepays, 
                           title = "Human biting rate by study area", 
                           log2 = TRUE)

ggplotly(plot)



a<- trmetrics_entomo %>%
    filter(genre == "Anopheles") %>%
  group_by(idpointdecapture,postedecapture) %>%
  summarise(ma=sum(ma)) %>%
  right_join(csh_all) %>%
  mutate(ma=ifelse(is.na(ma),0,ma)) %>%
  left_join(mean_date_by_mission) 




# ggplot(a, aes(x=date, y=ma,  colour = codepays, group = interaction(codepays, date))) + 
#   geom_boxplot(width=30) +
#       scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
#    scale_y_continuous(trans = 'log2', labels = scales::number_format(accuracy = 0.01))

 ggplot(a %>% filter(codepays=="BF"), aes(x=date, y=ma, group=date)) + 
   geom_boxplot(width=30) +
       scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  # scale_y_continuous(trans = 'log2', labels = scales::number_format(accuracy = 0.01)) +
   #facet_grid(rows = vars(codepays)) +
       theme_minimal() +
    labs(y = "ma") + 
    scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
    geom_vline(xintercept = as.numeric(mean(dates_intervention$start_intervention_date)),colour = "red", size = 0.3) +
    #annotate(geom="text",label="start of interventions",x=mean(dates_intervention$start_intervention_date)-10, y=mean(ma_b_country$val), colour = "red", angle = 90, size = 3.5) +
    #annotate(geom="text",label="Vertical red line = start of vector control interventions", colour = "red", size = 3.5, y=max(df$val), x=min(df$date)+160) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))



```

### By study area and by Anopheles complex, all intervention types

Average number of bites by Anopheles mosquitoes received per human per night, by Anopheles complex, split by study area.

`r emo::ji("warning")` : **y axis is log2 scaled**

```{r ma_b_country_cp}

df_n_obs <- csh_all %>%
  group_by(codepays,nummission) %>%
  summarise(n_obs = n())

ma <- fun_agg_data_entomo(trmetrics_entomo, df_n_obs, codepays, nummission, complex)  %>%
  group_split(codepays)
  
plot <- map(ma, ~fun_plot_ts_entomo(df = ., 
                            indicator_code = "ma", 
                            aes_color = complex, 
                            #facet = codepays, 
                           title = paste0("Human biting rate by anopheles complex, study area = ",unique(.$codepays)), 
                            log2 = TRUE))

ggplotly(plot[[1]])
ggplotly(plot[[2]])
```

### By study area and by intervention type, all anopheles species

`r emo::ji("warning")` : **y axis is log2 scaled**

```{r ma_b_country_interv}

df_n_obs <- csh_all %>%
  group_by(codepays,nummission,intervention) %>%
  summarise(n_obs = n())

ma <- fun_agg_data_entomo(trmetrics_entomo, df_n_obs, codepays, nummission, intervention) %>%
  group_split(codepays)

plot <- map(ma, ~fun_plot_ts_entomo(df = ., 
                           indicator_code = "ma", 
                           aes_color = intervention, 
                           title = paste0("Human biting rate by intervention type, study area = ",unique(.$codepays)), 
                           log2 = TRUE))

ggplotly(plot[[1]])
ggplotly(plot[[2]])








```




### By country and by anopheles species

Average number of bites by Anopheles mosquitoes received per human per night, by Anopheles species, split by study area.

**Note** : species was determined using qPCR techniques. Because of the large number of mosquitoes that were caught, not all the mosquitoes could be analysed. The table above the plot provides the number and proportion of mosquitoes that were analysed for species determination. In the plot, Ma was raised.

```{r ma_b_country_sp, eval=F}

ma_b_country_sp <- trmetrics_entomo %>%
  filter(genre == "Anopheles",!is.na(species)) %>%
  group_by(codepays,nummission,species) %>%
  summarise(val = sum(ma)) %>%
  left_join(df_n_obs) %>%
  mutate(val = round(val/n_obs,2)) %>%
  left_join(mean_date_by_mission) %>%
  dplyr::select(-n_obs) %>%
  as_tibble()

df_ratio_identified <- trmetrics_entomo %>% 
  filter(genre == "Anopheles") %>%
  mutate(species=ifelse(is.na(species),"unidentified","identified")) %>%
  group_by(codepays,nummission,species) %>%
  summarise(count = sum(ma)) %>%
  pivot_wider(names_from = species, values_from = count) %>%
  mutate(ratio_identified = round(identified / (identified + unidentified)*100))

# extrapolate
ma_b_country_sp <- ma_b_country_sp %>%
  left_join(df_ratio_identified) %>%
  mutate(val = 100 * val / ratio_identified)

b <- ggplot(ma_b_country_sp, aes(x = date, y = val)) +
    geom_line(aes(color = species), size = 0.5, linetype = "dashed") +
    geom_point(aes(color = species), size = 2) +
    theme_minimal() +
   labs(y = "ma (log2 scale)") + 
  scale_y_continuous(trans = 'log2') +
    scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  geom_vline(xintercept = as.numeric(mean(dates_intervention$start_intervention_date)),colour = "red", size = 0.3) +
 # annotate(geom="text",label="start of interventions",x=mean(dates_intervention$start_intervention_date)-10, y=mean(ma_b_country_sp$val), colour = "red", angle = 90, size = 3) +
    annotate(geom="text",label="Vertical red line = Start of vector control interventions", colour = "red", size = 3.5, y=max(ma_b_country$val), x=min(ma_b_country$date)+160) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  facet_grid(codepays ~ .)  +
  ggtitle("Human biting rate, by anopheles species, split by study area\n(extrapolated data)")

```
