---
title: "transmission_metrics"
author: "Paul Taconet"
date: "24/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) #set t
```

```{r packages}
library(ggplot2)
library(scales)
library(sf)
library(DBI)
library(RSQLite)
library(dplyr)
library(tidyr)
library(purrr)
library(rlang)
library(patchwork)
library(plotly)
library(htmlwidgets)
#library(kable)
library(kableExtra)
library(mapview)
library(stringr)
library(emo)
library(facetscales)
```

```{r db_connect}
# connect to database
path_to_db <- "data/react_db/react_db.gpkg"
react_gpkg <- DBI::dbConnect(RSQLite::SQLite(),dbname = path_to_db)

# open tables
trmetrics_entomo <- dbReadTable(react_gpkg, 'trmetrics_entomo')
entomo_csh_metadata_l1 <- dbReadTable(react_gpkg, 'entomo_csh_metadata_l1')
recensement_villages_l1 <- st_read(path_to_db,"recensement_villages_l1", quiet=T) %>%
  filter(!is.na(intervention)) %>%
  dplyr::select("codevillage","codepays","nomvillage","population","intervention","date_debut_interv","date_fin_interv","X","Y")

# # join to trmetrics_entomo the tables related to the HLC dates and locations + villages
# trmetrics_entomo <- trmetrics_entomo %>%
#   left_join(entomo_csh_metadata_l1[c("idpointdecapture","nummission","codevillage","pointdecapture","codepays","date_capture","X","Y")], by = "idpointdecapture") %>%
#   left_join(recensement_villages_l1[c("codevillage","intervention","date_debut_interv","date_fin_interv")], by = "codevillage") %>%
#   dplyr::select(-fid)

# get the number of human - nights
idpointdecaptures <- dbReadTable(react_gpkg, 'entomo_csh_metadata_l1') %>%
  dplyr::select(idpointdecapture,nummission,codevillage,pointdecapture,codepays,date_capture)

csh_exterior <- data.frame(idpointdecaptures, postedecapture = "e")
csh_interior <- data.frame(idpointdecaptures, postedecapture = "i")
csh_all <- rbind(csh_exterior,csh_interior) %>%
  left_join(recensement_villages_l1) %>%
  mutate(phase_interv = ifelse(as.Date(date_capture) < as.Date(date_debut_interv),"pre-intervention","post-intervention")) %>%
  mutate(phase_interv = ifelse((is.na(phase_interv) & codepays=="BF" & as.Date(date_capture) < as.Date("2017-08-17")),"pre-intervention",phase_interv)) %>%
  mutate(phase_interv = ifelse((is.na(phase_interv) & codepays=="CI" & as.Date(date_capture) < as.Date("2017-09-01")),"pre-intervention",phase_interv)) %>%
  mutate(phase_interv = ifelse((is.na(phase_interv) & codepays=="BF" & as.Date(date_capture) > as.Date("2017-08-17")),"post-intervention",phase_interv)) %>%
  mutate(phase_interv = ifelse((is.na(phase_interv) & codepays=="CI" & as.Date(date_capture) > as.Date("2017-09-01")),"post-intervention",phase_interv))


# get mean date for each entomo mission in each country
mean_date_by_mission <- entomo_csh_metadata_l1 %>%
  dplyr::select(-geom) %>%
  group_by(codepays,nummission) %>%
  summarise(date = mean(as.Date(date_capture))) %>%
  as_tibble()

# get start and end date of intervention for each country
dates_intervention <- recensement_villages_l1 %>%
  group_by(codepays) %>%
  summarise(start_intervention_date=min(as.Date(date_debut_interv),na.rm = T),end_intervention_date=max(as.Date(date_fin_interv),na.rm = T))
  
```

```{r rainbow_plots, eval=F}
#source("https://raw.githubusercontent.com/RainCloudPlots/RainCloudPlots/master/tutorial_R/R_rainclouds.R")
source("https://raw.githubusercontent.com/RainCloudPlots/RainCloudPlots/master/tutorial_R/summarySE.R")
#source("https://raw.githubusercontent.com/RainCloudPlots/RainCloudPlots/master/tutorial_R/simulateData.R")
library(cowplot)
library(forcats)

#see https://wellcomeopenresearch.org/articles/4-63
geom_flat_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                             position = "dodge", trim = TRUE, scale = "area",
                             show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomFlatViolin,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      trim = trim,
      scale = scale,
      ...
    )
  )
}

#' @rdname ggplot2-ggproto
#' @format NULL
#' @usage NULL
#' @export
GeomFlatViolin <-
  ggproto("GeomFlatViolin", Geom,
    setup_data = function(data, params) {
      data$width <- data$width %||%
        params$width %||% (resolution(data$x, FALSE) * 0.9)

      # ymin, ymax, xmin, and xmax define the bounding rectangle for each group
      data %>%
        group_by(group) %>%
        mutate(
          ymin = min(y),
          ymax = max(y),
          xmin = x,
          xmax = x + width / 2
        )
    },

    draw_group = function(data, panel_scales, coord) {
      # Find the points for the line to go all the way around
      data <- transform(data,
        xminv = x,
        xmaxv = x + violinwidth * (xmax - x)
      )

      # Make sure it's sorted properly to draw the outline
      newdata <- rbind(
        plyr::arrange(transform(data, x = xminv), y),
        plyr::arrange(transform(data, x = xmaxv), -y)
      )

      # Close the polygon: set first and last point the same
      # Needed for coord_polar and such
      newdata <- rbind(newdata, newdata[1, ])

      ggplot2:::ggname("geom_flat_violin", GeomPolygon$draw_panel(newdata, panel_scales, coord))
    },

    draw_key = draw_key_polygon,

    default_aes = aes(
      weight = 1, colour = "grey20", fill = "white", size = 0.5,
      alpha = NA, linetype = "solid"
    ),

    libraryd_aes = c("x", "y")
  )


a <- trmetrics_entomo %>%
    filter(genre == "Anopheles") %>%
  group_by(idpointdecapture, postedecapture) %>%
  summarise(ma = sum(val)) %>%
  right_join(csh_all) %>%
  mutate(ma = ifelse(is.na(ma), 0, ma)) %>%
  left_join(mean_date_by_mission) %>%
  as_tibble()



p9 <- ggplot(a %>% filter(codepays=="BF"),aes(x = forcats::fct_rev(phase_interv), y = ma, fill = forcats::fct_rev(phase_interv), colour = forcats::fct_rev(phase_interv)))+
  geom_flat_violin(position = position_nudge(x = .25, y = -1), adjust =2, trim = TRUE)+
  geom_point(position = position_jitter(width = .15), size = .25)+
  geom_boxplot(aes(x = forcats::fct_rev(phase_interv), y = ma),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +
  ylab('Score')+xlab('Group')+theme_cowplot()+guides(fill = FALSE, colour = FALSE) +
  facet_wrap(~intervention)+
  scale_colour_brewer(palette = "Dark2")+
  scale_fill_brewer(palette = "Dark2")+
  scale_y_continuous(limits = c(-1,NA)) +
  ggtitle("Figure R9: Complex Raincloud Plots with Facet Wrap")

p9


```

```{r functions, eval=F}

fun_agg_data_entomo <- function(trmetrics_entomo, df_n_obs, var, ...){
  
  group_var <- enquos(...)
  
  df_metric <- trmetrics_entomo %>%
    filter(genre == "Anopheles", var == var) %>%
    group_by(!!! group_var) %>%
    summarise(val = sum(val)) %>%
    left_join(df_n_obs) %>%
    mutate(val = round(val/n_obs,2)) %>%
    left_join(mean_date_by_mission) %>%
    dplyr::select(-n_obs) %>%
    as_tibble()
  
  return(df_metric)

}

fun_plot_ts_entomo <- function(df, indicator_code, aes_color, facet = NULL, title, log2 = FALSE){
  
  aes_color <- enquo(aes_color)
  facet <- enquo(facet)

    th_plot <- ggplot(df, aes(x = date, y = val)) + 
    geom_line(aes(color = !! aes_color), size = 0.5, linetype = "dashed") +
    geom_point(aes(color = !! aes_color), size = 2) +
    #scale_color_manual(values = c("#00AFBB", "#E7B800")) +
    theme_minimal() +
    labs(y = indicator_code) + 
    scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
    #ylim(-0.01,NA) +
    geom_vline(xintercept = as.numeric(mean(dates_intervention$start_intervention_date)),colour = "red", size = 0.3) +
    #annotate(geom="text",label="start of interventions",x=mean(dates_intervention$start_intervention_date)-10, y=mean(ma_b_country$val), colour = "red", angle = 90, size = 3.5) +
    #annotate(geom="text",label="Vertical red line = start of vector control interventions", colour = "red", size = 3.5, y=max(df$val), x=min(df$date)+160) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle(title)
  
  if(log2){
    th_plot <- th_plot + scale_y_continuous(trans = 'log2', limits = c(-0.2,NA), labels = scales::number_format(accuracy = 0.01)) + labs(y = paste0(indicator_code," (log2 scale)"))
  }
   if(!quo_is_null(facet)){
     th_plot <- th_plot + facet_grid(rows = vars(!!facet))
   }

      return(th_plot)
}


```

Click a village to get additional information

```{r map }
areas <- st_read(path_to_db,"contexte_frontieresreact", quiet=T)
mapview(areas, alpha.regions = 0.1, aplha = 1, layer.name = "Study areas") + mapview(recensement_villages_l1, zcol = "intervention", cex = 4, alpha.regions = 0.9, aplha = 1, layer.name = "Villages")
```

Following indicators are provided for each study area : 

```{r indicators }
googlesheets4::sheets_deauth()
df_indicators <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1dIeSOa2WinXvOQGLmIjA0gFdsHnb6zMMsDME-G5pyMc/edit?usp=sharing", sheet = "var_reponse3", col_types="c")
df_indicators %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```


##  Ma : Human biting rate 

### All Anopheles species

Number of bites by Anopheles mosquitoes received per human per night, by study area

```{r ma_b_country}
# by country

ma_an <- trmetrics_entomo %>%
  filter(genre=="Anopheles", var=="ma", as.numeric(nummission) <= 8) %>%
  group_by(idpostedecapture,codepays, nummission, intervention) %>%
  summarise(val = sum(val))

bf <- ggplot(ma_an %>% filter(codepays == "BF", nummission <= 8), aes(x = nummission, y = val, fill = intervention)) +
   geom_boxplot(outlier.shape = NA) +
   ylim(0,40) + # 40 for BF, 400 for CI
   geom_vline(xintercept = 3.5, colour = "red", size = 0.3) +  # 3.4 for BF, 4.5 for CI
   annotate(geom="text",label="start of interventions",x = 3.7, y = 37, colour = "red", angle = 90, size = 3.5) +
   theme_minimal() +
   labs(y = "Human biting rate (all Anoph. species)") 


ci <- ggplot(ma_an %>% filter(codepays == "CI", nummission <= 8), aes(x = nummission, y = val, fill = intervention)) +
   geom_boxplot(outlier.shape = NA) +
   ylim(0,400) + # 40 for BF, 400 for CI
   geom_vline(xintercept = 4.5, colour = "red", size = 0.3) +  # 3.4 for BF, 4.5 for CI
   annotate(geom="text",label="start of interventions",x = 4.7, y = 370, colour = "red", angle = 90, size = 3.5) +
   theme_minimal() +
   labs(y = "Human biting rate (all Anoph. species)") 

# 
# 
# scales_y <- list(
#   BF = scale_y_continuous(limits = c(0,40)),
#   CI = scale_y_continuous(limits = c(0,400))
# )
# 
# pl <- ggplot(ma_an %>% filter(nummission <= 8), aes(x = as.character(nummission), y = val, fill = intervention)) +
#    geom_boxplot(outlier.shape = NA) +
#   #facet_wrap(~codepays, scale="free",  dir = "v") +
#     facet_grid_sc(rows = vars(codepays), scales = list(y = scales_y)) +
#    #ylim(0,400) + # 40 for BF, 400 for CI
#    geom_vline(xintercept = 4.5, colour = "red", size = 0.3) +  # 3.4 for BF, 4.5 for CI
#    annotate(geom="text",label="start of interventions",x = 4.7, y = 370, colour = "red", angle = 90, size = 3.5) +
#    theme_minimal() +
#    labs(y = "Human biting rate (all Anoph. species)") 

bf
ci

```

### By major Anopheles species

Number of bites by Anopheles mosquitoes received per human per night, by Anopheles species, split by study area.

**Note** : species was determined using qPCR techniques. Because of the large number of mosquitoes that were caught, not all the mosquitoes could be analysed. The table provides the number and proportion of mosquitoes that were analysed for species determination. In the plot, Ma was raised.

```{r ma_b_country_sp}

ma_b_country_sp <- trmetrics_entomo %>%
  filter(genre == "Anopheles",var == "ma", !is.na(species)) %>%
  group_by(idpostedecapture,codepays,nummission,species,intervention) %>%
  summarise(val = sum(val))

df_ratio_identified <- trmetrics_entomo %>% 
  filter(genre == "Anopheles",var == "ma") %>%
  mutate(species=ifelse(is.na(species),"unidentified","identified")) %>%
  group_by(idpostedecapture,species) %>%
  summarise(count = sum(val)) %>%
  pivot_wider(names_from = species, values_from = count, values_fill =  list(count = 0)) %>%
  mutate(ratio_identified = round(identified / (identified + unidentified)*100))

# extrapolate
ma_sp <- ma_b_country_sp %>%
  left_join(df_ratio_identified) %>%
  mutate(val = 100 * val / ratio_identified) %>%
  filter(species %in% c("An.coluzzii","An.funestus_ss","An.gambiae_ss")) %>%
  filter(as.numeric(nummission) <= 8)

# ma_sp <- trmetrics_entomo %>%
#   filter(genre=="Anopheles", var=="ma", as.numeric(nummission) <= 8) %>%
#   group_by(idpostedecapture, codepays, species, nummission, intervention) %>%
#   summarise(val = sum(val)) %>%
#   filter(species %in% c("An.coluzzii","An.funestus_ss","An.gambiae_ss"))


bf <- ggplot(ma_sp %>% filter(codepays == "BF", nummission <= 8), aes(x = nummission, y = val, fill = intervention)) +
   geom_boxplot(outlier.shape = NA) +
   ylim(0,20) + 
   geom_vline(xintercept = 3.5, colour = "red", size = 0.3) + 
   annotate(geom="text",label="start of interventions",x = 3.7, y = 37, colour = "red", angle = 90, size = 3.5) +
   labs(y = "Human biting rate") + 
  facet_wrap(~species, scales = "free", ncol = 1)

ci <- ggplot(ma_sp %>% filter(codepays == "CI", nummission <= 8), aes(x = nummission, y = val, fill = intervention)) +
   geom_boxplot(outlier.shape = NA) +
   #ylim(0,20) + # 40 for BF, 400 for CI
   geom_vline(xintercept = 4.5, colour = "red", size = 0.3) +  # 3.4 for BF, 4.5 for CI
   #annotate(geom="text",label="start of interventions",x = 4.7, y = 37, colour = "red", angle = 90, size = 4.5) +
   labs(y = "Human biting rate") + 
  facet_wrap(~species, scales = "free", ncol = 1)

bf
ci
```
